---
title: "Blue whiting Operating Model"
subtitle: "Beverton and Holt SRR with steepess fixed at 0.9"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    word_document:
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
tags: FLPKG FLR
license: Creative Commons Attribution-ShareAlike 4.0 International
---
# Notes


```{r knitr_init, echo=FALSE, results="hide", dpi=100}
library(knitr)
## Global options
opts_chunk$set(echo    =FALSE,
               eval    =TRUE,
               cache   =TRUE,
               cache.path="cache/sim1/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warning =FALSE,
               fig.height=4.5,
               fig.width =8,
               fig.path  ="tex/sim1-")

options(digits=3)
```
```{r, pkgs, echo=FALSE, message=FALSE}
library(FLCore)
library(ggplotFL)
library(FLfse)
library(kobe)
library(TMB)
library(stockassessment)
library(FLBRP)
library(FLasher)

library(plyr)
library(reshape)
library(magrittr)

theme_set(theme_bw(16))
```
```{r, dirs, echo=FALSE, message=FALSE}
dirMy =dirname(dirname(FLife:::getScriptPath())) 
#dirMy=getwd()
dirDat=file.path(dirMy,"data")   
```
```{r}
source(file.path(dirMy,"Rmd/lFig.R"))  
iFig =25   
```
```{r}
refpts=FLPar(c(
  blim    =1500000,
  bpa     =2250000,
  flim    =0.88,
  fpa     =0.53,
  msytrig =2250000,
  fmsy    =0.32,
  bmgtlow =1500000,
  bmgt    =2250000,
  fmgstlow=0.05,
  fmgt    =0.32),units="NA")
```

```{r, smry}
load(file.path(dirDat,"om.RData")) 
load(file.path(dirDat,"sims.RData"))

nms       =c("HCR1","HCR2","HCR1 Bounded","HCR2 Bounded","Historical","Fmsy")
names(nms)=c("sim1","sim2","sim1_bnd","sim2_bnd","Historical","fmsy")
```

```{r, fig.height=8,fig.width=8, dpi=100}
plot(FLStocks("Historical"=om1,
              "Fmsy"      =iter(sims[["fmsy"]][[1]],1),
              "HCR(Fmsy)" =iter(sims[["sim0.0"]][[1]],1)))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("",values=rainbow(10)[c(1,9,7)])
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


```{r, fig.height=8,fig.width=8, dpi=100}
plot(FLStocks("Fmsy"=sims[["fmsy"]][[1]],
              "HCR(Fmsy)"=sims[["sim0.0"]][[1]],
              "HCR(Fmsy) with Ass. Error"=sims[["sim0"]][[1]]))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rainbow(3))
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 

\newpage
\blandscape

```{r, fig.height=8,fig.width=12, dpi=100}
p=plot(FLStocks("HCR I" =sims[["sim1"]][[1]],
                "HCR II"=sims[["sim2"]][[1]]))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rainbow(3))

p2=plot(FLStocks("HCR I" =iter(sims[["sim1"]][[1]],1),
                 "HCR II"=iter(sims[["sim2"]][[1]],1)))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rep("black",2))

p+geom_line(aes(date,data),data=p2$data,linetype=2)+
  facet_grid(qname~stock,scale="free")+
  scale_color_manual(values=rep("black",2))+
  scale_fill_manual(values=rep("grey",2))+
  theme(legend.position="none")+xlab("Year")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 

\newpage
```{r, fig.height=8,fig.width=12, dpi=100}
p=plot(FLStocks("HCR I"        =sims[["sim1"]][[1]],  
                "HCR I bounded"=sims[["sim1_bnd"]][[1]]))+ 
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rainbow(3))

p2=plot(FLStocks("HCR I"        =iter(sims[["sim1"]][[1]],1),  
                 "HCR I bounded"=iter(sims[["sim1_bnd"]][[1]],1)))+ 
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rep("black",2))

p+geom_line(aes(date,data),data=p2$data,linetype=2)+
  facet_grid(qname~stock,scale="free")+
  scale_color_manual(values=rep("black",2))+
  scale_fill_manual(values=rep("grey",2))+
  theme(legend.position="none")+xlab("Year")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


\newpage
```{r, fig.height=8,fig.width=12, dpi=100}
p=plot(FLStocks("HCR II"        =sims[["sim2"]][[1]],  
                "HCR II bounded"=sims[["sim2_bnd"]][[1]]))+ 
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rainbow(3))

p2=plot(FLStocks("HCR II"        =iter(sims[["sim2"]][[1]],1),  
                 "HCR II bounded"=iter(sims[["sim2_bnd"]][[1]],1)))+ 
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rep("black",2))

p+geom_line(aes(date,data),data=p2$data,linetype=2)+
  facet_grid(qname~stock,scale="free")+
  scale_color_manual(values=rep("black",2))+
  scale_fill_manual(values=rep("grey",2))+
  theme(legend.position="none")+xlab("Year")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 

```{r}
#Time series with quantiles for stock, recruitment, catch, ... 

### Summary stats
series<-function(object,start,end){
  res=FLQuants(window(object,start=start,end=end),
               catch  =catch,
               ssb    =ssb,
               biomass=computeStock,
               juve   =function(x) computeStock(x)-ssb(x),
               esb    =function(x) apply((stock.n(x)%*%stock.wt(x))%/%
                                         (harvest(x)%/%fbar(x)),c(2,6),sum),
               rec    =rec,
               f      =fbar)
  res}

pts=ldply(sims[c("sim1","sim2","sim1_bnd","sim2_bnd","Historical","fmsy")],function(x)
    as.data.frame(series(x[[1]],2000,2018),drop=TRUE))

pts$qname=factor(pts$qname,
                 levels=c("rec","ssb","biomass","juve","f","catch","esb"),
                 labels=c("Rec","SSB","Biomass","juvenile","F","Catch","Exploitable \nBiomass"))

qtl=ddply(pts,.(.id,qname,year), with,
                    t(as.data.frame(quantile(data,c(0.025,0.25,0.5,0.75,0.975)),drop=TRUE)))

aav=ddply(subset(pts,year>=2000),.(.id,qname,iter), with,
                data.frame(aav=mean(abs((data[-1]-data[-length(data)])/data[-1]))))

chk=ldply(sims[c("sim1","sim2","sim1_bnd","sim2_bnd")], function(x) x[[2]]) 

tc=ddply(subset(pts,qname=="Catch"),.(.id,iter), with,
         data.frame(sum=sum(data),var=sd(data)^0.5,
                    aav=c(mean(abs((data[-1]-data[-length(data)])/data[-1])))))

save(pts,qtl,aav,chk,tc,file=file.path(dirDat,"smry1.RData"),compress="xz")
```

\newpage
```{r, ts1, fig.height=8, fig.width=12, dpi=100}
qtl$Scenario=factor(qtl$.id,labels=nms,
                            levels=names(nms))
pts$Scenario=factor(pts$.id,labels=nms,
                            levels=names(nms))

ggplot(subset(qtl,qname%in%c("Catch","SSB","F","Exploitable \nBiomass")))+
  geom_ribbon(aes(year,ymin=`25%`, ymax=`75%`),alpha=.5,fill="red")+
  geom_ribbon(aes(year,ymin=`2.5%`,ymax=`97.5%`),alpha=.1,fill="red")+
  geom_line(  aes(year,`50%`))+
  geom_line(  aes(year,data),col="blue",size=0.3,
              data=subset(pts,qname%in%c("Catch","SSB","F","Exploitable \nBiomass")&iter==3))+
  facet_grid(qname~Scenario,scale="free")+
  theme_bw(16)+
  xlab("Year")+ylab("")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 
\elandscape

```{r, chk1, fig.height=8,fig.width=8, dpi=100}
par1=array(c(ftar=refpts["fmsy"],btrig=refpts["msytrig"],fmin=FLPar(0.05),blim=refpts["blim"]),
           dim=c(4),
           dimnames=list(c("ftar","btrig","fmin","blim")))

par2=array(c(0.20,2250000,
             0.05,1500000,
             0.32,5200000,
             0.20,4000000),
          dim=c(4,2),
          dimnames=list(c("ftar","btrig","fmin","blim"),c("lower","upper")))[c(1,3,2,4),1:2]

ggplot(subset(chk,.id%in%c("sim1","sim2")))+   
  geom_vline(aes(xintercept=c(par2[c("btrig"),"upper"])),linetype=2,size=.2)+
  geom_vline(aes(xintercept=c(par2[c("btrig"),"lower"])),linetype=2,size=.2)+
  geom_vline(aes(xintercept=c(par2[c("blim"), "upper"])),linetype=2,size=.2)+
  geom_vline(aes(xintercept=c(par2[c("blim"), "lower"])),linetype=2,size=.2)+
  geom_hline(aes(yintercept=c(par2[c("ftar"), "lower"])),linetype=2,size=.2)+
  geom_hline(aes(yintercept=c(par2[c("ftar"), "upper"])),linetype=2,size=.2)+
  geom_hline(aes(yintercept=c(par1[c("fmin")])),linetype=2,size=.2)+
  geom_point(aes(ssb,f),col="red")+
  expand_limits(x=0, y=0)+
  scale_x_continuous(limits=c(0,0.75e7))+
  facet_grid(.id~.)+
  theme_bw(16)+
  xlab("SSB")+ylab("F")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


```{r, dpi=100}
library(grid)
source('/home/laurence/Desktop/flr/kobe/R/kobe-phaseMar.R')
source('/home/laurence/Desktop/flr/kobe/R/kobe-phase.R')

bref=c(par2[c("btrig"),"upper"],
       par2[c("btrig"),"lower"],
       par2[c("blim"), "upper"],
       par2[c("blim"), "lower"])

fref=c(par2[c("ftar"), "lower"],
       par2[c("ftar"), "upper"],     
       par1[c("fmin")])

pts=ldply(sims,function(x)
  model.frame(FLQuants(window(x[[1]],start=2000,end=2018),                "stock"=ssb,"harvest"=fbar),drop=T))[,c(4:5,1,2)]
```

```{r, fig.height=10, fig.width=10, dpi=100}
kobePhase(subset(transform(pts,stock=stock/1000),.id=="sim0.0"),ylim=c(0,1),
          quadcol=rep(c("white"),4))+ 
  facet_wrap(~year)+
  geom_point(aes(ssb/1000,f),col="red",size=.1,
           data=subset(chk,.id%in%c("sim1"))[,c("ssb","f")])+ 
  geom_point(aes(stock,harvest),size=.3,fill="cyan",colour="blue",shape=23)+
  xlab("SSB")+ylab("F")+
  scale_y_continuous(limit=c(0,1))+
  theme_bw(16)+
  theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```

**Figure `r iFig  =iFig  +1; iFig  `** `r lFig[iFig  ]`  

```{r, fig.height=10, fig.width=10, dpi=100}
kobePhase(subset(transform(pts,stock=stock/1000),.id=="sim1"),ylim=c(0,1),
          quadcol=rep(c("white"),4))+ 
  facet_wrap(~year)+
  geom_point(aes(ssb/1000,f),col="red",size=.1,
           data=subset(chk,.id%in%c("sim1"))[,c("ssb","f")])+
  geom_point(aes(stock,harvest),size=.3,fill="cyan",colour="blue",shape=23)+
  xlab("SSB")+ylab("F")+
  scale_y_continuous(limit=c(0,1))+
  theme_bw(16)+
  theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]`  

```{r, fig.height=10, fig.width=10, dpi=100}
kobePhase(subset(transform(pts,stock=stock/1000),.id=="sim2"),ylim=c(0,1),
          quadcol=rep(c("white"),4))+
  facet_wrap(~year)+
  geom_point(aes(ssb/1000,f),col="red",size=.1,
           data=subset(chk,.id%in%c("sim2"))[,c("ssb","f")])+
  geom_point(aes(stock,harvest),size=.3,fill="cyan",colour="blue",shape=23)+
  xlab("SSB")+ylab("F")+
  scale_y_continuous(limit=c(0,1))+
  theme_bw(16)+
  theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]`

```{r, fig.height=8, fig.width=8, dpi=100}
library(grid)
source('/home/laurence/Desktop/flr/kobe/R/kobe-phase.R')
source('/home/laurence/Desktop/flr/kobe/R/kobe-phaseMar.R')

bref=c(par2[c("btrig"),"upper"],
       par2[c("btrig"),"lower"],
       par2[c("blim"), "upper"],
       par2[c("blim"), "lower"])

fref=c(par2[c("ftar"), "lower"],
       par2[c("ftar"), "upper"],     
       par1[c("fmin")])

kobePhaseMar2(subset(pts,year==2012&.id=="sim1"),xlim=c(0,1e7),ylim=c(0,1),
              quadcol=rep(c("white"),4),
              bref=par1["blim"],fref=par1["ftar"],layer=
geom_point(aes(ssb,f),col="red",size=.2,
           data=subset(chk,.id%in%c("sim1"))), 
xlab="SSB",ylab="F",fourth=NULL)
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]`  


```{r, fig.height=8, fig.width=8, dpi=100}
kobePhaseMar2(subset(pts,year==2012&.id=="sim2"),xlim=c(0,1e7),ylim=c(0,1),quadcol=rep(c("white"),4),
              bref=par1["blim"],fref=par1["ftar"],layer=
geom_point(aes(ssb,f),col="red",size=.2,
           data=subset(chk,.id%in%c("sim2"))), 
xlab="SSB",ylab="F",fourth=NULL)
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]`  

```{r, tc, fig.width=8,fig.height=8, dpi=100}
dodge=position_dodge(width=1)

nms       =c("HCR1","HCR2","HCR1 Bounded","HCR2 Bounded","Historical","Fmsy")
names(nms)=c("sim1","sim2","sim1_bnd","sim2_bnd","Historical","fmsy")


ggplot(melt(tc,id=c(".id","iter")),aes(x=nms[.id], y=value))+
    geom_violin(position=dodge)+
    geom_boxplot(width=0.3, position=dodge,outlier.shape=NA)+
    scale_fill_manual(values=c("white", "white"))+
    facet_grid(variable~.,scale="free")+
    theme_bw(16)+
  xlab("Scenario")+ylab("")+
  theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


```{r, aav1, fig.height=10, fig.width=8, dpi=100}
aav$Scenario=factor(aav$.id,labels=nms,
                            levels=names(nms))
ggplot(aav,aes(x=Scenario, y=aav))+
    geom_violin(position=dodge)+
    geom_boxplot(width=0.3, position=dodge,outlier.shape=NA)+
    scale_fill_manual(values=c("white", "white"))+
  facet_grid(qname~.,scale="free")+
  theme_bw(16)+
  xlab("Scenario")+ylab("")+
  theme(axis.text.x=element_text(angle=-30),legend.position="bottom")

```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 

```{r, stab1, dpi=100}
dat<-chk%>% 
      subset(ssb>c(refpts["bpa"])&.id%in%c("sim1","sim2","sim1_bnd","sim2_bnd"))%>% 
      with(table(.id,hcrYrs))%>%
      melt

dat$Scenario=factor(dat$.id,labels=nms,
                            levels=names(nms))

scaleFUN <- function(x) paste(sprintf("%.0f", x/10),"%",sep="")

ggplot(dat)+
  geom_line(aes(hcrYrs,value,col=nms[.id]))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_y_continuous(labels=scaleFUN)+
                       #scales::percent_format(accuracy=0.1))+
  xlab("Year")+ylab("Stability Mechanism Applied")+
  scale_color_manual("Scenario",values=rainbow(4))
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 

```{r, fig.height=5, echo=FALSE, dpi=100}
pts=ldply(sims,function(x)
  model.frame(FLQuants(window(x[[1]],start=2000,end=2018),                "stock"=ssb,"harvest"=fbar),drop=T))
names(pts)[1]="run"
pts$bpa =pts$stock<2.25e6
pts$blim=pts$stock<1.5e6

ggplot(ddply(subset(pts,!(run%in%c("0.6","ICES","fmsy",
                                   "sim0","sim0.0","sim0_bnd",
                                   "acf1","acf2"))),
                    .(run,year), with,mean(bpa)))+
  geom_line(aes(year,V1,col=nms[run]))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  xlab("Year")+ylab("P(SSB<Bpa)")+
  scale_color_manual("",values=rainbow(8))+
  scale_color_manual("Scenario",values=rainbow(5))
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


```{r, fig.height=5, echo=FALSE, dpi=100}
pts=ldply(sims,function(x)
  model.frame(FLQuants(window(x[[1]],start=2000,end=2018),                "stock"=ssb,"harvest"=fbar),drop=T))
names(pts)[1]="run"
pts$bpa =pts$stock<2.25e6
pts$blim=pts$stock<1.5e6

ggplot(ddply(subset(pts,!(run%in%c("0.6","ICES","fmsy",
                                   "sim0","sim0.0","sim0_bnd",
                                   "acf1","acf2"))),
                    .(run,year), with,mean(blim)))+
  geom_line(aes(year,V1,col=nms[run]))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  xlab("Year")+ylab("P(SSB<Blim)")+
  scale_color_manual("",values=rainbow(8))+
  scale_color_manual("Scenario",values=rainbow(5))
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


```{r, fig.height=10,fig.width=8, dpi=100}
load(file.path(dirDat,"escape.RData"))

plot(FLStocks(sims[["Historical"]][[1]],"Escapement"=escape))+
  theme_bw()
```

**Figure `r iFig =iFig +1; iFig `** `r lFig[iFig ]` 


```{r, tab1, eval=FALSE, dpi=100}
hcr1=subset(chk,.id%in%c("sim1","sim1_bnd","sim1"))  
with(cbind(hcr1,cut=cut(hcr1$ssb,c(0,par1["blim"],par1["btrig"],Inf))),table(.id,cut))

hcr2=subset(chk,.id%in%c("sim2","sim2_bnd","sim2"))
with(cbind(hcr1,cut=cut(hcr2$ssb,c(0,par2["blim","lower"],par2["btrig","lower"],par2["blim","upper"],par2["btrig","upper"],Inf))),table(.id,cut))
```

