---
title: "Blue whiting HCR Evaluation"
subtitle: "Operating Model Conditioning"
author: "Laurence Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    word_document:
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
tags: FLPKG FLR
license: Creative Commons Attribution-ShareAlike 4.0 International
---
# Note


```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(echo    =FALSE,
               eval    =TRUE,
               cache   =TRUE,
               cache.path="cache/om/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warning =FALSE,
               fig.height=4.5,
               fig.width =8,
               fig.path  ="tex/om-")

options(digits=3)

iFig=4
```
```{r, pkgs, echo=FALSE, message=FALSE}
library(FLCore)
library(ggplotFL)
library(FLfse)
library(kobe)
library(TMB)
library(stockassessment)
library(FLBRP)
library(FLasher)

library(plyr)
library(reshape)
library(magrittr)

theme_set(theme_bw(16))
```
```{r, dirs, echo=FALSE, message=FALSE}
#dirMy =dirname(dirname(FLife:::getScriptPath()))
dirMy ="/home/laurence/Desktop/sea++/PELAG/bluewhiting"
dirDat="/home/laurence/Desktop/Dropbox/bluewhiting/data"
```
```{r}
source(file.path(dirMy,"Rmd/lFig.R"))
```
```{r}
refpts=FLPar(c(
  blim    =1500000,
  bpa     =2250000,
  flim    =0.88,
  fpa     =0.53,
  msytrig =2250000,
  fmsy    =0.32,
  bmgtlow =1500000,
  bmgt    =2250000,
  fmgstlow=0.05,
  fmgt    =0.32),units="NA")
```

```{r, om2, eval=FALSE}
load(file.path(dirDat,"samData.RData"))

conf=loadConf(dat,"/home/laurence/Desktop/sea++/PELAG/bluewhiting/sam/BW_2018/conf/model.cfg")
par =defpar(dat,conf)
fit =sam.fit(dat,conf,par)

om =SAM2FLStock(fit)
err=SAM_uncertainty(fit, n=1000, print_screen=FALSE, idx_cov=FALSE)

om1=om
stock.n(om)=err[["stock.n"]]
harvest(om)=err[["harvest"]]
units(harvest(om))="f"

save(om,om1, file=file.path(dirDat,"om.RData"),compress="xz")
```
```{r, retro, eval=FALSE}
load(file.path(dirDat,"samModel.RData"))
load(file.path(dirDat,"om.RData"))
retro=retro(fit, year=10)

retro=FLStocks(llply(retro,
          function(x){ 
              f=exp(x[[2]][[15]])
              n=exp(x[[2]][[16]])
              res=om1[,seq(dim(n)[2])]
              stock.n(res)[]=n
              harvest(res)[seq(dim(f)[1]),]=f
              harvest(res)[dim(f)[1]+1,]   =harvest(res)[dim(f)[1],]
            
              res}))

names(retro)=2019-seq(10)
save(retro, file=file.path(dirDat,"retro.RData"),compress="xz")
```

```{r, fig.height=8, dpi=100}
load(file.path(dirDat,"om.RData"))

d=plot(om)
d$data$qname=factor(d$data$qname,levels=c("Catch","Rec","SSB","F"))
d+theme_bw(16)+theme(legend.position="none")+
  xlab("Year")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, fig.width=6,fig.height=8, dpi=100}
dat=as.data.frame(FLQuants(om1,
                  stock.wt=stock.wt,
                  catch.wt=catch.wt,
                  m=m,
                  mat=mat,
                  sel=function(x)harvest(om1)%/%fbar(om1)),drop=T)
dat$qname=factor(dat$qname,levels=c("stock.wt","catch.wt","mat","m","sel"),
                           labels=c("Stock Mass",
                                    "Catch Mass",
                                    "Maturity",
                                    "M",
                                    "Selection"))
ggplot(data=dat)+
  geom_line(aes(age,data,col=ac(year),group=year))+
  geom_smooth(aes(age,data))+
  facet_wrap(~qname,scale="free",ncol=2)+
  xlab("Age")+ylab("")+
  theme_bw(16)+
  theme(legend.position="none")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, fig.width=8,fig.height=8, dpi=100}
ggplot(data=subset(dat,!(qname%in%c("M","Maturity","Catch Mass"))))+
  geom_line(aes(year,data,group=age,col=ac(age)))+
  facet_wrap(~qname,scale="free",ncol=1)+
  xlab("Year")+ylab("")+
  theme_bw(18)+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(10))
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, stkn, fig.height=4, fig.width=8, dpi=100}
dat=stock.n(om1)%/%apply(stock.n(om1),1,mean)
ggplot(as.data.frame(dat))+
  geom_point(aes(year,age,size=data))+
  theme_bw(16)+
  theme(legend.position="none")+
  xlab("Year")+ylab("Age")
```

**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]`


```{r, ctcn, fig.height=4, fig.width=8, dpi=100}
dat=catch.n(om1)%/%apply(catch.n(om1),1,mean)
ggplot(as.data.frame(dat))+
  geom_point(aes(year,age,size=data))+
  theme_bw(16)+
  theme(legend.position="none")+
  xlab("Year")+ylab("Age")
```

**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]`

```{r, fig.height=6, dpi=100}
ccf(c(rec(om1)),c(ssb(om1)),main="")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, fig.height=6, dpi=100}
esb    =function(x) apply((stock.n(x)%*%stock.wt(x))%/%
                                         (harvest(x)%/%fbar(x)),c(2,6),sum)
ccf(c(rec(om1)),c(esb(om1)),main="")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, sg, fig.height=5, fig.width=6, dpi=100}
sg=fmle(as.FLSR(om1,model="segreg"),control=list(silent=TRUE))
eqlSG=FLBRP(om1,sr=sg)  
fbar(eqlSG)=fbar(eqlSG)/(12)
refpts(eqlSG)=refpts(eqlSG)[c(1,2,3)]

dat=model.frame(FLQuants(rec    =fitted(sg), 
                            ssb    =ssb(sg),
                            rec.obs=rec(sg)),drop=T)

ggplot(dat)+
  geom_line( aes(ssb,rec))+  
  geom_path( aes(ssb,rec.obs),linetype=2)+
  geom_text( aes(ssb,rec.obs,label=year),data=subset(dat,year%in%c(1980,2018,1990,2000,2001)))+  
  geom_point(aes(ssb,rec.obs),col="red")+
  theme_bw(16)+
  xlab("SSB")+ylab("Recruitment")
```

**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 
                                         
```{r, bh, fig.height=5, fig.width=6, dpi=100}
sr=fmle(as.FLSR(om1,model="bevholt"),control=list(silent=TRUE))
eql=FLBRP(om1,sr=sr)   
 
dat=model.frame(FLQuants(rec    =fitted(sr), 
                         ssb    =ssb(sr),
                         rec.obs=rec(sr)),drop=T)
ggplot(dat)+
  geom_line( aes(ssb,rec))+  
  geom_path( aes(ssb,rec.obs),linetype=2)+
  geom_text( aes(ssb,rec.obs,label=year),data=subset(dat,year%in%c(1980,2018,1990,2000,2001)))+  
  geom_point(aes(ssb,rec.obs),col="red")+
  theme_bw(16)+
  xlab("SSB")+ylab("Recruitment")
```

**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, bhsv, fig.height=5, fig.width=6, dpi=100}
sr9=fmle(as.FLSR(om1,model="bevholtSV"),
                fixed=c(s=0.9,spr0=spr0(eql)),
                        control=list(silent=TRUE))
eql9=FLBRP(om1,sr=ab(sr9)) 
refpts(eql9)=refpts(eql9)[c(1,2,3)]

dat=model.frame(FLQuants(rec    =fitted(sr9), 
                            ssb    =ssb(sr9),
                            rec.obs=rec(sr9)),drop=T)

ggplot(dat)+
  geom_line( aes(ssb,rec))+  
  geom_path( aes(ssb,rec.obs),linetype=2)+
  geom_text( aes(ssb,rec.obs,label=year),data=subset(dat,year%in%c(1980,2018,1990,2000,2001)))+  
  geom_point(aes(ssb,rec.obs),col="red")+
  theme_bw(16)+
  xlab("SSB")+ylab("Recruitment")
```

**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, bh-regimes, dpi=100}
ggplot(as.data.frame(residuals(sr),drop=TRUE))+
      geom_point(aes(year,data),col="grey25")+
      geom_line( aes(year,data),col="grey75")+
    geom_polygon(aes(year,data,group=regime),
          fill="lavender",col="blue",
          lwd=.25,alpha=.5,
          data=FLife:::rod(residuals(sr)))+
      xlab("Year")+ylab("Residual")+
      theme_bw(16)+
      theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, sg-regimes, dpi=100}
ggplot(as.data.frame(residuals(sg),drop=TRUE)) +
      geom_point(aes(year,data),col="grey25")+
      geom_line( aes(year,data),col="grey75")+
    geom_polygon(aes(year,data,group=regime),
          fill="lavender",col="blue",
          lwd=.25,alpha=.5,
          data=FLife:::rod(residuals(sg)))+
      xlab("Year")+ylab("Residual")+
      theme_bw(16)+
      theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, bh9-regimes, dpi=100}
ggplot(as.data.frame(residuals(sr9),drop=TRUE))+
      geom_point(aes(year,data),col="grey25")+
      geom_line( aes(year,data),col="grey75")+
    geom_polygon(aes(year,data,group=regime),
          fill="lavender",col="blue",
          lwd=.25,alpha=.5,
          data=FLife:::rod(residuals(sr9)))+
      xlab("Year")+ylab("Residual")+
      theme_bw(16)+
      theme(axis.text.x=element_text(angle=-30),legend.position="bottom")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, bh-acf, fig.height=6, dpi=100}
acf(iter(residuals(sr9),1),main="Recruit Deviates")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, acf-sim, dpi=100}
set.seed(67979)
ggplot(as.data.frame(rlnoise(2,rec(om1)%=%0,0.6,0.8)))+
  geom_line(aes(year,data,col=iter))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  xlab("Year")+ylab("")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, sg-eql, fig.height=6, fig.width=6, dpi=100}
plot(eqlSG,refpts=FALSE)+
  theme_bw(16)
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, bh-eql, fig.height=6, fig.width=8, dpi=100}
refpts(eql)=refpts(eql)[c(1,2,3)]
plot(eql)+
  theme_bw(16)
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, bh-eql9, fig.height=6, fig.width=8, dpi=100}
plot(eql9)+
  theme_bw(16)
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 


```{r, fig.width=8,fig.height=10, dpi=100}
dat=mdply(as.numeric(dimnames(m(om1))$year[-seq(3)]),function(year){
      x =window(om1,start=year-5,end=year)
      sr=fmle(as.FLSR(x,model="geomean"),control=list(silent=TRUE))
      eq=FLBRP(x,sr=sr)
      as.data.frame(refpts(eq)[c("f0.1"),1:4])})

dat$quant=factor(dat$quant,levels=c("rec","ssb","yield","harvest"),
                           labels=c("Recruits","Bmsy","MSY","Fmsy")) 
dat[dat$quant=="Recruits","data"][]=c(rec(om1)[,-(seq(5))])

ggplot(dat)+
  geom_line(aes(as.numeric(X1)+1979,data,col=refpt))+
  facet_grid(quant~.,scale="free")+
  xlab("Year")+ylab("")+
  scale_color_manual("",values=rainbow(3))+
  theme_bw(16)+
  theme(legend.position="none")

```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r}
save(om,om1,sr,sr9,eql,eql9,file=file.path(dirDat,"om.RData"),compress="xz")
```

```{r, retro-plot, fig.height=8, dpi=100}
load(file.path(dirDat,"retro.RData"))

plot(retro)+
  theme_bw(16)+
  theme(legend.position="bottom")+
  xlab("Year")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 

```{r, ass-err}
assEr=laply(retro, function(x){
  yr=ac(dims(fbar(x))$maxyear)
  
  c(f  =c(catch(x  )[,yr]/computeStock(x  )[,yr])/
        c(catch(om1)[,yr]/computeStock(om1)[,yr]),
    ssb=c(ssb(x)[,yr])/c(ssb(om1)[,yr]))})

save(assEr, file=file.path(dirDat,"assEr.RData"),compress="xz")
```

```{r, ass-err-plot, fig.height=10, fig.width=10, dpi=100}
library(grid)
source('/home/laurence/Desktop/flr/kobe/R/kobe-phaseMar.R')
source('/home/laurence/Desktop/flr/kobe/R/kobe-phase.R')

dt2=as.data.frame(rmvnorm(1000,c(1,1),cov(assEr)))

names(dt2)=c("stock","harvest")

kobePhaseMar(dt2,quadcol=c("white","white","white","white"),
  xlab="F",ylab="SSB")
```
**Figure `r iFig=iFig+1;iFig`** `r  lFig[iFig]` 



