---
title: "Blue whiting HCR Evaluation"
author: "Laurence Kell & Polina Levontin"
date: "July 29, 2019"
output: ioslides_presentation
---

```{r knitr_init, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
## Global options
opts_chunk$set(echo    =FALSE,
               eval    =TRUE,
               cache   =TRUE,
               cache.path="cache/stakeholder/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warning =FALSE,
               fig.height=4.5,
               fig.width =8,
               fig.path  ="tex/stakeholder-")

options(digits=3)
```

```{r}
library(FLCore)
library(ggplotFL)
library(FLBRP)
library(plyr)
library(kobe)
library(grid)
library(png)
library(reshape)

dirDat="/home/laurence/Desktop/Dropbox/bluewhiting/data"   
load(file.path(dirDat,"om.RData"))
load(file.path(dirDat,"sims.RData"))
load(file.path(dirDat,"smry1.RData"))
load(file.path(dirDat,"retro.RData"))

pts=cast(subset(pts,qname%in%c("SSB","F")),.id+year+iter~qname,value="data")[,c(2,4,5,1)]
names(pts)=c("year","stock","harvest","run")

attributes(pts)[[2]]=c("data.frame")
```

## Outline

+ Objectives
+ Approach
+ Example Results
+ Summary
+ Questions
+ Next Steps

# Objectives
## Objectives

### Evaluate alternative HCRs that can achieve objectives for both 
+ sustainability and 
+ economic factors

### Compare alternative HCRs to historical outcomes using a hindcast

# Resource Dynamics 

## Large variability in assessment results between years

```{r, retro, fig.height=5, fig.width=6}
p=plot(FLStocks(llply(retro,function(x) window(x,start=2001))))+
  theme_bw(16)+
  theme(legend.position="none")
p$data=subset(p$data,qname!="Catch")
p
```

## Variable recruitment 

```{r}
plot(rec(om1)/1e3)+
  theme_bw(16)+xlab("Year")+ylab("Recruits (1000s)")
```

## No clear stock recruitment relationship
```{r, fig.height=5, echo=FALSE}
sr=fmle(as.FLSR(om1,model="bevholt"),control=list(silent=TRUE))
eql=FLBRP(om1,sr=sr)   
 
dat=model.frame(FLQuants(rec    =fitted(sr), 
                         ssb    =ssb(sr),
                         rec.obs=rec(sr)),drop=T)
ggplot(dat)+
  geom_line( aes(ssb,rec))+  
  geom_path( aes(ssb,rec.obs),linetype=2)+
  geom_text( aes(ssb,rec.obs,label=year),data=subset(dat,year%in%c(1980,2018,1990,2000,2001)))+  
  geom_point(aes(ssb,rec.obs),col="red")+
  theme_bw(16)+
  xlab("SSB")+ylab("Recruitment")
```


## Year-class strength

```{r, yrcl, fig.height=4, fig.width=8}
dat=stock.n(om1)%/%apply(stock.n(om1),1,mean)
ggplot(as.data.frame(dat))+
  geom_point(aes(year,age,size=data))+
  theme_bw(16)+
  theme(legend.position="none")+
  xlab("Year")+ylab("Age")
```

## Selection Pattern

```{r}
ggplot(as.data.frame((harvest(om1)%/%fbar(om1))[,c("2001","2010","2018")]))+
  geom_line(aes(age,data,col=ac(year)))+
  theme_bw(20)+xlab("Age")+ylab("Selectivity")+
  scale_colour_manual("Year",values=rainbow(3))
```

## FMSY

```{r, fig.height=5, echo=FALSE}
dat=mdply(as.numeric(dimnames(m(om1))$year[-seq(3)]),function(year){
      x =window(om1,start=year-5,end=year)
      sr=fmle(as.FLSR(x,model="geomean"),control=list(silent=TRUE))
      eq=FLBRP(x,sr=sr)
      as.data.frame(refpts(eq)[c("f0.1"),1:4])
      })

dat$quant=factor(dat$quant,levels=c("rec","ssb","yield","harvest"),
                           labels=c("Recruits","Bmsy","MSY","Fmsy")) 
dat[dat$quant=="Recruits","data"][]=c(rec(om1)[,-(seq(5))])

ggplot(dat)+
  geom_line(aes(as.numeric(X1)+1979,data,col=refpt),size=2)+
  facet_grid(quant~.,scale="free")+
  xlab("Year")+ylab("")+
  scale_color_manual("",values=rainbow(3))+
  theme_bw(16)+
  theme(legend.position="none")
```

## Assumptions

### Assessment Error of 30% on SSB estimate
### Stock Recruitment relationship 
- Beverton and Holt wth steepness=0.9
- if SSB~0 then there will be no recruitment
- variability in recruitment taken from the residual in the historical fit

### Growth, selection pattern in the future is known

```{r}
refpts=FLPar(c(
  blim    =1500000,
  bpa     =2250000,
  flim    =0.88,
  fpa     =0.53,
  msytrig =2250000,
  fmsy    =0.32,
  bmgtlow =1500000,
  bmgt    =2250000,
  fmgstlow=0.05,
  fmgt    =0.32),units="NA")

par1=array(c(ftar=refpts["fmsy"],btrig=refpts["msytrig"],fmin=FLPar(0.05),blim=refpts["blim"]),
           dim=c(4),
           dimnames=list(c("ftar","btrig","fmin","blim")))

par2=array(c(0.20,2250000,
             0.05,1500000,
             0.32,5200000,
             0.20,4000000),
          dim=c(4,2),
          dimnames=list(c("ftar","btrig","fmin","blim"),c("lower","upper")))[c(1,3,2,4),1:2]

bref=c(par2[c("btrig"),"upper"],
       par2[c("btrig"),"lower"],
       par2[c("blim"), "upper"],
       par2[c("blim"), "lower"])

fref=c(par2[c("ftar"), "lower"],
       par2[c("ftar"), "upper"],     
       par1[c("fmin")])
```

# HCRs

## Projection at $F_{MSY}$  

```{r, fig.height=5}
plot(FLStocks("Historical"=om1,
              "Fmsy"      =iter(sims[["fmsy"]][[1]],1)))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("",values=rainbow(10)[c(1,9,7)])
```

## Projection at $F_{MSY}$ and HCR with $F_{MSY}$ target

```{r, fig.height=5}
plot(FLStocks("Historical"=om1,
              "Fmsy"      =iter(sims[["fmsy"]][[1]],1),
              "HCR(Fmsy)" =iter(sims[["sim0.0"]][[1]],1)))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("",values=rainbow(10)[c(1,9,7)])
```

## HCR with assessment error

```{r, fig.height=5}
plot(FLStocks("Fmsy"=sims[["fmsy"]][[1]],
              "HCR(Fmsy) with Ass. Error"=sims[["sim0"]][[1]],
              "Single Realisation"=iter(sims[["sim0"]][[1]],1)))+
  theme_bw(16)+
  theme(legend.position="bottom")+
  scale_colour_manual("Management",values=rainbow(4)[-3])
```


```{r}
fn1<-function(yr) {
  ymax=max(c(ldply(sims[c("sim1")], function(x)
     as.data.frame(quantile(c(catch(x[[1]])[,ac(2001:2018)]),0.95),drop=T))[,2]))
  
  fourth=ggplot(ldply(sims[c("sim1")], function(x)
     as.data.frame(catch(x[[1]])[,ac(yr)],drop=T)))+geom_violin(aes(.id,data,fill=.id))+
  theme(legend.position="none")+
    scale_fill_manual(values=colorRampPalette(c("orange","blue"),space="Lab")(2))+
    scale_y_continuous(limits=c(0,ymax))
                       
  kobePhaseMar2(subset(pts,year==yr&run=="sim1"),xlim=c(0,1e7),ylim=c(0,1),
              quadcol=rep(c("white"),4),
              bref=2250000,fref=0.32,
              layer=geom_point(aes(ssb,f),col="red",size=.2,
                               data=subset(chk,.id%in%c("sim1","sim1_bnd"))), 
              xlab="SSB",ylab="F",fourth=fourth)}   
   
fnBnd<-function(yr) {  
  ymax=max(c(ldply(sims[c("sim1","sim1_bnd")], function(x)
     as.data.frame(quantile(c(catch(x[[1]])[,ac(2001:2018)]),0.95),drop=T))[,2]))

    fourth=ggplot(ldply(sims[c("sim1","sim1_bnd")], function(x)
     as.data.frame(catch(x[[1]])[,ac(yr)],drop=T)))+geom_violin(aes(.id,data,fill=.id))+
  theme(legend.position="none")+
    scale_fill_manual(values=colorRampPalette(c("orange","blue"),space="Lab")(2))+
    scale_y_continuous(limits=c(0,ymax))

    kobePhaseMar2(subset(pts,year==yr&
            run%in%c("sim1","sim1_bnd")),
            xlim=c(0,1e7),ylim=c(0,1),
              quadcol=rep(c("white"),4),
              bref=2250000,fref=0.32,layer=
geom_point(aes(ssb,f),col="red",size=.2,
           data=subset(chk,.id%in%c("sim1","sim1_bnd"))), 
xlab="SSB",ylab="F",fourth=fourth)}  
```

# Simulations
## HCR I:   2001 
```{r, hcr1_2001}
fn1(2001)    
```

## HCR I:   2002 
```{r, hcr1_2002}
fn1(2002) 
```

## HCR I:   2003 
```{r, hcr1_2003}
fn1(2003) 
```

## HCR I:   2004 
```{r, hcr1_2004}
fn1(2004) 
```

## HCR I:   2005 
```{r, hcr1_2005}
fn1(2005) 
```

## HCR I:   2006 
```{r, hcr1_2006}
fn1(2006) 
```

## HCR I:   2007 
```{r, hcr1_2007}
fn1(2007) 
```

## HCR I:   2008 
```{r, hcr1_2008}
fn1(2008)
```

## HCR I:   2009 
```{r, hcr1_2009}
fn1(2009) 
```

## Year-class strength

```{r, stkn, fig.height=4, fig.width=8}
dat=stock.n(om1)%/%apply(stock.n(om1),1,mean)
ggplot(as.data.frame(dat))+
  geom_point(aes(year,age,size=data))+
  theme_bw(16)+
  theme(legend.position="none")+
  xlab("Year")+ylab("Age")
```

```{r,eval=FALSE}
plot(rec(om1)/1e3)+
  theme_bw()+xlab("Year")+ylab("Recruits (1000s)")
```

## HCR I:   2009 
```{r}
fn1(2009) 
```

## HCR I:   2010 
```{r, hcr1_2010}
fn1(2010) 
```

## HCR I:   2011 
```{r, hcr1_2011}
fn1(2011) 
```

## HCR I:   2012
```{r, hcr1_2012}
fn1(2012) 
```

## HCR I:   2013 
```{r, hcr1_2013}
fn1(2013) 
```

## HCR I:   2014 
```{r, hcr1_2014}
fn1(2014) 
```

## HCR I:   2015 
```{r, hcr1_2015}
fn1(2015) 
```

## HCR I:   2016 
```{r, hcr1_2016}
fn1(2016) 
```

## HCR I:   2017 
```{r, hcr1_2017}
fn1(2017) 
```

## HCR I:   2018 
```{r, hcr1_2018}
fn1(2018) 
```

# Bounds

## Bounds:   2001 
```{r, hcr1bnd_2001}
fnBnd(2001)   
```

## Bounds:   2002 
```{r, hcr1bnd_2002}
fnBnd(2002) 
```

## Bounds:   2003 
```{r, hcr1bnd_2003}
fnBnd(2003) 
```

## Bounds:   2004 
```{r, hcr1bnd_2004}
fnBnd(2004) 
```

## Bounds:   2005 
```{r, hcr1bnd_2005}
fnBnd(2005) 
```

## Bounds:   2006 
```{r, hcr1bnd_2006}
fnBnd(2006) 
```

## Bounds:   2007 
```{r, hcr1bnd_2007}
fnBnd(2007) 
```

## Bounds:   2008 
```{r, hcr1bnd_2008}
fnBnd(2008) 
```

## Bounds:   2009 
```{r, hcr1bnd_2009}
fnBnd(2009) 
```

## Bounds:   2010 
```{r, hcr1bnd_2010}
fnBnd(2010) 
```

## Bounds:   2011 
```{r, hcr1bnd_2011}
fnBnd(2011) 
```

## Bounds:   2012 
```{r, hcr1bnd_2012}
fnBnd(2012) 
```

## Bounds:   2013 
```{r, hcr1bnd_2013}
fnBnd(2013) 
```

## Bounds:   2014 
```{r, hcr1bnd_2014}
fnBnd(2014) 
```

## Bounds:   2015 
```{r, hcr1bnd_2015}
fnBnd(2015) 
```

## Bounds:   2016 
```{r, hcr1bnd_2016}
fnBnd(2016) 
```

## Bounds:   2017 
```{r, hcr1bnd_2017}
fnBnd(2017) 
```

## Bounds:   2018 
```{r, hcr1bnd_2018}
fnBnd(2018) 
```

```{r}
fn12<-function(yr) {
    ymax=max(c(ldply(sims[c("sim1","sim2")], function(x)
     as.data.frame(quantile(c(catch(x[[1]])[,ac(2001:2018)]),0.95),drop=T))[,2]))

  fourth=ggplot(ldply(sims[c("sim1","sim2")], function(x)
     as.data.frame(catch(x[[1]])[,ac(yr)],drop=T)))+geom_violin(aes(.id,data,fill=.id))+
  theme(legend.position="none")+
    scale_fill_manual(values=colorRampPalette(c("orange","blue"),space="Lab")(2))+
    scale_y_continuous(limits=c(0,ymax))
                                           
    kobePhaseMar2(subset(pts,year==yr&
            run%in%c("sim1","sim2")),
            xlim=c(0,1e7),ylim=c(0,1),
              quadcol=rep(c("white"),4),
              bref=2250000,fref=0.32,layer=
geom_point(aes(ssb,f,col=.id),size=.2,
           data=subset(transform(chk,.id=as.character(.id)),.id%in%c("sim1","sim2"))), 
xlab="SSB",ylab="F",fourth=fourth)
}  
```

# HCR I v II 

## HCR I v II:   2001 
```{r, hcr12_2001}
fn12(2001) 
```

## HCR I v II:   2002 
```{r, hcr12_2002}
fn12(2002) 
```

## HCR I v II:   2003 
```{r, hcr12_2003}
fn12(2003) 
```

## HCR I v II:   2004 
```{r, hcr12_2004}
fn12(2004) 
```

## HCR I v II:   2005 
```{r, hcr12_2005}
fn12(2005) 
```

## HCR I v II:   2006 
```{r, hcr12_2006}
fn12(2006) 
```

## HCR I v II:   2007 
```{r, hcr12_2007}
fn12(2007) 
```

## HCR I v II:   2008 
```{r, hcr12_2008}
fn12(2008) 
```

## HCR I v II:   2009 
```{r, hcr12_2009}
fn12(2009) 
```

## HCR I v II:   2010 
```{r, hcr12_2010}
fn12(2010) 
```

## HCR I v II:   2011 
```{r, hcr12_2011}
fn12(2011) 
```

## HCR I v II:   2012 
```{r, hcr12_2012}
fn12(2012) 
```


## HCR I v II:   2013 
```{r, hcr12_2013}
fn12(2013) 
```

## HCR I v II:   2014 
```{r, hcr12_2014}
fn12(2014) 
```

## HCR I v II:   2015 
```{r, hcr12_2015}
fn12(2015) 
```

## HCR I v II:   2016 
```{r, hcr12_2016}
fn12(2016) 
```

## HCR I v II:   2017 
```{r, hcr12_2017}
fn12(2017) 
```

## HCR I v II:   2018 
```{r, hcr12_2018}
fn12(2018) 
```


## Summary Statistics

# Summary Statistics

## Total Catch
```{r, fig.height=5, echo=FALSE}
grid.raster(readPNG("/home/laurence/Desktop/sea++/PELAG/bluewhiting/Rmd/tex/sim1-tc-1.png"))
```

## Summary Statistics
```{r, fig.height=5, echo=FALSE}
grid.raster(readPNG("/home/laurence/Desktop/sea++/PELAG/bluewhiting/Rmd/tex/sim1-aav1-1.png"))
```


## P(SSB<Bpa) in any year
```{r, fig.height=5, echo=FALSE}
pts=ldply(sims,function(x)
  model.frame(FLQuants(window(x[[1]],start=2000,end=2018),                "stock"=ssb,"harvest"=fbar),drop=T))
names(pts)[1]="run"
pts$bpa =pts$stock<2.25e6
pts$blim=pts$stock<1.5e6

ggplot(ddply(subset(pts,(run%in%c("sim1","sim2","sim1_bnd","sim2_bnd","fmsy","Historical"))),
                    .(run,year), with,mean(bpa)))+
  geom_line(aes(year,V1,col=run))+
  theme_bw()+
  theme(legend.position="bottom")+
  xlab("Year")+ylab("P(SSB<Bpa)")+
  scale_color_manual("",values=rainbow(8))
```

## P(SSB<Blim) in any year
```{r, fig.height=5, echo=FALSE}
pts=ldply(sims,function(x)
  model.frame(FLQuants(window(x[[1]],start=2000,end=2018),                "stock"=ssb,"harvest"=fbar),drop=T))
names(pts)[1]="run"
pts$bpa =pts$stock<2.25e6
pts$blim=pts$stock<1.5e6

ggplot(ddply(subset(pts,(run%in%c("sim1","sim1_bnd",
                                   "sim2","sim2_bnd",
                                   "Historical","fmsy"))),
                    .(run,year), with,mean(blim)))+
  geom_line(aes(year,V1,col=run))+
  theme_bw()+
  theme(legend.position="bottom")+
  xlab("Year")+ylab("P(SSB<Blim)")+
  scale_color_manual("",values=rainbow(8))
```

## Next Steps

+ OM Assumptions
+ Assessment Error
+ Alternative HCRs
+ Summary Statistics
